Carbonate chemistry at Ny-Ålesund

### 2021-05-06 awipev-CO2.Rmd
Looking into salinity data sent pby Philipp

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group_pressure = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & depth <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & depth <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & depth <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    ),
    depth_group_philipp = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
tmp <- d_hour %>% 
  dplyr::select(date, depth, pressure_insitu_ctd) %>% 
  dplyr::mutate(month=month(d_hour$date, label = TRUE)) %>% 
  group_by(month)
depth_labels <- c("0-2", "2-4", "4-6", "6-8", "8-12")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::mutate(depth_group = depth_group_pressure) %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)

>>>>

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))

d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
# read a file provided by Philipp on 2021-04-09 which fills most gaps of pressure data
# depth_complete replaces all other depth and pressure variables
svl_remos_depth <- tibble(
  readRDS(file = paste0(
    path,
    "fb_data/svl_remos_depth.rds"
  )))
d_hour <- dplyr::left_join(d_hour, svl_remos_depth, by = 'datetime') %>% 
  dplyr::select(-adcp_fill, -ctd_fill) %>% 
  dplyr::rename(depth = depth_complete)
# check_plot <- d_hour %>% 
#   ggplot() + geom_point(aes(x=datetime, y=depth, col='blue', alpha = 0.7)) +
#   geom_point(aes(x=datetime, y=depth, col='red', alpha = 0.3))

# # read a file provided by Philipp on 2021-04-13 in the hope that it contains
# # in situ CTD salinity data for any month of June. IT DOES NOT
# sal_2021_04_13 <-
#   readRDS(file = paste0(
#     path,
#     "fb_data/remos_2021-04-13/svl_remos_ctd_salinity_2017_2020.rds"
#   )) %>% 
#   dplyr::rename(sal_new = svl_ctd_salinity_mean)
# zz <- dplyr::left_join(d_hour, sal_2021_04_13, by = 'datetime') %>%
#   dplyr::mutate(month = month(date)) %>%
#   dplyr::filter(depth < 4, month(date) == 6, datetime > as.POSIXct("2017-07-01 11:00:00"))
# pp <- ggplot(zz, aes(x=datetime, y=sal_new)) +
#     geom_point()

d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth_group_pressure = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & depth <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & depth <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & depth <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 10, NA)
          )
        )
      )
    ),
    depth_group_philipp = ifelse( # create new variable depth
      depth <= 2,
      1,
      ifelse(
        depth > 2 & depth <= 4,
        3,
        ifelse(
          depth > 4 & depth <= 6,
          5,
          ifelse(
            depth > 6 & depth <= 8,
            7,
            ifelse(depth > 8, 10, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
rm(d_hour)
```

@@

## Stratification
```{r stratification, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
tmp <- d_all %>% 
  dplyr::select(date, depth, pressure_insitu_ctd) %>% 
  dplyr::mutate(month=month(d_all$date, label = TRUE)) %>% 
  group_by(month)
depth_labels <- c("0-2", "2-4", "4-6", "6-8", "8-12")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::mutate(depth_group = depth_group_pressure) %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
